---
title: "STL Lead Inequity"
author: "IMS"
format: html
editor_options: 
  chunk_output_type: console
---

## Attach the tidyverse and janitor packages in a new code chunk

```{r}
library(tidyverse)
library(janitor)
```

## Read in the stl_blood_lead.csv data as stl_lead and use janitor::clean_names to convert all variable names to lower snake case

```{r}
stl_blood_lead <- read_csv(here::here("data", "stl_blood_lead.csv"))

stl_blood_lead <- clean_names(stl_blood_lead)
names(stl_blood_lead)
```

## Do some basic exploration of the dataset (e.g. using summary, data visualizations and summary statistics).

```{r}
skimr::skim(stl_blood_lead)
```

```{r}
stl_blood_lead_plot <- ggplot(data = stl_blood_lead,
                        aes(x = poverty_u18, y = pct_elevated)) +
  geom_point() +
  theme_minimal() +
  labs(title = "Toolik Station Chlorophyll a concentration",
       x = "Poverty Under 18",
       y = "Percent Lead Elevated Blood")
stl_blood_lead_plot
```


# In a new code chunk,
- from `stl_blood_lead` create a new data frame called `stl_lead_prop` that has one additional column called `prop_white` that returns the percent of each census tract identifying as white (variable `white` in the dataset divided by variable `totalPop`, times 100). You may need to do some Googling. Hint: d`plyr::mutate(new_col = col_a / col_b)` will create a new column new_col that contains the value of col_a / col_b

```{r}
stl_lead_prop <- stl_blood_lead
stl_lead_prop <-  stl_lead_prop %>%  #loading data 
  dplyr::mutate(prop_white = (white / total_pop)*100)

```

# Create a scatterplot
- In a new code chunk, create a scatterplot graph of the percentage of children in each census tract with elevated blood lead levels (pctElevated) versus the percent of each census tract identifying as white.
- Customize by updating several aesthetics (e.g. size, opacity (see alpha =), color, etc.)
- Store the scatterplot as stl_lead_plot
- Have the scatterplot returned in the knitted html - customize the size that it appears when knitted
- Also save a .png of the scatterplot to figs, with dimensions of (6” x 5”) (width x height)
- In text above or below the scatterplot, write 1 - 2 sentences describing the overall trend that you observe from your graph


```{r}
stl_lead_plot <- ggplot(stl_lead_prop, 
       aes(x=white, y = pct_elevated))+
  geom_point(aes(color=prop_white))+
  theme_minimal() +
  labs(x="Total White Population", y="Percent Elevated")
  
stl_lead_plot

ggsave(here::here("figs", "stl_lead_plot.png"), height = 6, width = 5)
```

As The total white population increases, the percent elevated lead in blood decreases.

# Create a histogram
- Create a histogram of only the pctElevated column in the data frame (remember, this will only take one variable - the frequency is calculated for you by geom_histogram)
- Customize the fill, color, and size aesthetics - test some stuff! Feel free to make it awful.
- Once you’ve played around with customization, export the histogram as a .jpg to the figs folder
- Make sure the histogram also shows up in your rendered html

```{r}
stl_lead_hist <- ggplot(stl_lead_prop, aes(x=pct_elevated))+
  geom_histogram(fill = "white", color = "grey", binwidth = 5) +
  labs(x="Percent Elevated Lead in Blood", y="Count", title = "Blood Lead Hist")
  
stl_lead_hist

ggsave(here::here("figs", "stl_lead_hist.png"), height = 6, width = 5)
```

Entry 1: Ian
- In a punk band
- from Ventura county
- ......... to be continued